---
title: "Short Video Platform Analytics"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
runtime: shiny
---

```{r global, include=FALSE, message=TRUE, warning=TRUE}
# Load necessary packages
library(dplyr)
library(DT)
library(flexdashboard)
library(ggplot2)
library(ggsci)
library(plotly)
library(readr)
library(scales)
library(shiny)
library(stringr)
library(tibble)

# Load dataset (setwd() to repo/lab04)
df <- read_csv("main_chinese_data.csv")
categories_cn_en <- read_csv("categories_cn_en.csv")

#####

# load the full dataset (setwd() to repo/lab04)
interaction_filtered <- read_csv("interaction_filtered_small.csv")

#####

# Define main city tiers for mapping
tiers_main <- c("first-tier city", "new first-tier city", "second-tier city", "third-tier city")
df$plot_tier <- ifelse(df$fre_city_level_en %in% tiers_main, df$fre_city_level_en, "other")

# Convert character engagement columns to logical (if needed) and sum up engagement
engage_cols <- c("click", "liked.", "follow", "collect", "forward")
df[engage_cols] <- lapply(df[engage_cols], function(x) as.logical(ifelse(x == "True", TRUE, FALSE)))
df$engagement <- rowSums(df[engage_cols], na.rm = TRUE)

# Function to get hourly engagement data
get_hourly_engagement <- function(hour) {
  hour <- as.numeric(hour)
  df %>%
    filter(p_hour == hour) %>%
    group_by(plot_tier) %>%
    summarise(
      total_engagement = sum(engagement, na.rm = TRUE),
      avg_engagement = mean(engagement, na.rm = TRUE),
      count = n(),
      .groups = 'drop'
    ) %>%
    arrange(desc(total_engagement))
}

# Function to get all hours data for time series
get_all_hours_data <- function() {
  df %>%
    group_by(p_hour, plot_tier) %>%
    summarise(
      total_engagement = sum(engagement, na.rm = TRUE),
      avg_engagement = mean(engagement, na.rm = TRUE),
      .groups = 'drop'
    )
}
```

Overview
========

Help, about, logo etc.

Dataset from the paper ["A Large-scale Dataset with Behavior, Attributes, and Content of Mobile Short-video Platform"](https://arxiv.org/pdf/2502.05922)
(https://github.com/tsinghua-fib-lab/ShortVideo_dataset)

Categories {data-orientation=rows}
==========

```{r }
category_data <- reactive(interaction_filtered %>%
    filter(category_id %in% input$categories) %>%
    # add category name from categories_cn_en
    inner_join(categories_cn_en, by = "category_id") %>%
    # use category_name_en as category_name
    mutate(category_name = category_name_en)
)

category_engagement = function() { category_data() %>%
  summarise(
    views = n(),
    clicks = sum(click, na.rm = TRUE),
    comments = sum(comment, na.rm = TRUE),
    follows = sum(follow, na.rm = TRUE),
    collects = sum(collect, na.rm = TRUE),
    forwards = sum(forward, na.rm = TRUE),
    hates = sum(hate, na.rm = TRUE),
    likes = sum(cvm_like, na.rm = TRUE),
    .groups = 'drop'
  )
}

ALPHA = 0.6

# generate a color map from category_name that we will use for every
# plot (we want a shared legend)
# (Use the default ggplot theme, because others won't get applied
# for some reason)
category_colors <- reactive({
  if (nrow(category_data()) == 0) {
    return(NA)
  }
  gg_color_hue <- function(n) {
    hues = seq(15, 375, length = n + 1)
    hcl(h = hues, l = 65, c = 100)[1:n]
  }
  setNames(
    gg_color_hue(length(unique(category_data()$category_name))),
    unique(category_data()$category_name)
  )
})

# create a ggplot theme with category_colors()
# (Somehow make this reactive)

theme_category <- function() {
  return (labs(color=NULL, alpha=NULL, fill=NULL) +
    scale_alpha_continuous(range = c(ALPHA, ALPHA), guide="none")
  )
}

PLOTLY = function (plot) {
  # tooltip: don't show alpha, color, fill
  ggplotly(plot, tooltip = c("x", "y")) %>%
    hide_legend()
} 
```

Column {.sidebar data-width=250}
-----------------------------------------------------------------------

```{r results = "asis"}
# only lvl 1 categories
tv_data <- categories_cn_en %>%
  filter(category_level == 1) %>%
  # First letter always uppercase
  mutate(title = str_to_title(category_name_en)) %>%
  arrange(category_name_en)

# create c(title = category_id)
logger::log_info(paste0("category_id: ", toString(tv_data$category_id)))
logger::log_info(paste0("title: ", toString(tv_data$title)))
tv_data_c <- setNames(tv_data$category_id, tv_data$title)

# combo box
renderUI({
  selectInput("categories", "Categories:", tv_data_c, multiple=T)
})
```

```{r}
# And HERE, show legend, common for all plots. (show in alphabetical
# order) (HTML)

renderUI({
  colors = category_colors()
  
  if (any(is.na(colors))) {
    return(tags$div(style="color: red", "Please select a category."))
  }

  colors <- colors[order(names(colors))]
  tags$div(
    style = "margin: 10px;",
    lapply(names(colors), function(name) {
      # small colored square + label
      tags$div(
        style = paste0("display: flex; align-items: center; margin-bottom: 5px;"),
        tags$div(
          style = paste0("width: 20px; height: 20px; flex: 0 0 20px; background-color: ", colors[name], "; margin-right: 10px; border-radius: 50%;"),
        ),
        tags$span(name)
      )
    })
  )
})
```

Row
-----------------------------------------------------------------------

### Views

```{r results = "asis"}
renderValueBox({ valueBox(format(category_engagement()$views), color = "primary") })
```

### Clicks

```{r results = "asis"}
renderValueBox({ valueBox(format(category_engagement()$clicks), color = "primary") })
```

### Comments

```{r results = "asis"}
renderValueBox({ valueBox(format(category_engagement()$comments), color = "primary") })
```

### Follows

```{r results = "asis"}
renderValueBox({ valueBox(format(category_engagement()$follows), color = "primary") })
```

### Collects

```{r results = "asis"}
renderValueBox({ valueBox(format(category_engagement()$collects), color = "primary") })
```

### Forwards

```{r results = "asis"}
renderValueBox({ valueBox(format(category_engagement()$forwards), color = "primary") })
```

### Hates

```{r results = "asis"}
renderValueBox({ valueBox(format(category_engagement()$hates), color = "primary") })
```

### Likes

```{r results = "asis"}
renderValueBox({ valueBox(format(category_engagement()$likes), color = "primary") })
```

Row
-----------------------------------------------------------------------

### User's Gender

```{r results = "asis"}
renderPlotly({
  cd <- category_data() %>% group_by(user_id, gender)
  plot <- cd %>% ggplot(aes(group=category_name, fill=category_name, x=gender))
  if (nrow(cd) > 0) {
    plot <- plot +
      stat_count(aes(y = ..prop.., group = category_name, fill = category_name), position = "dodge") +
      labs(y=NULL, x=NULL) +
      scale_x_discrete(labels=c("M"="Male", "F"="Female")) +
      scale_color_manual(values = category_colors(), guide="none") +
      scale_fill_manual(values = category_colors(), guide="none") +
      theme_category()
  }
  PLOTLY(plot)
})
```

### User's Age

```{r results = "asis"}
renderPlotly({
  plot <- category_data() %>%
    ggplot(aes(group=category_name, color=category_name, fill=category_name, x=age)) +
    geom_density(aes(alpha=ALPHA)) +
    labs(y="density", x=NULL) +
    scale_color_manual(values = category_colors(), guide="none") +
    scale_fill_manual(values = category_colors(), guide="none") +
    theme_category()
  PLOTLY(plot)
})
```

### User's Phone Price

```{r results = "asis"}
renderPlotly({
  plot <- category_data() %>%
    ggplot(aes(group=category_name, color=category_name, fill=category_name, x=mod_price)) +
    geom_density(aes(alpha=ALPHA)) +
    labs(y="density", x="phone price ($)") +
    scale_color_manual(values = category_colors(), guide="none") +
    scale_fill_manual(values = category_colors(), guide="none") +
    theme_category()
  PLOTLY(plot)
})
```

Row
-----------------------------------------------------------------------

### Subcategories

```{r results = "asis"}
renderTable({
  # from categories_cn_en take category_id and category_name_en
  # of categories which have parent_id %in% input$categories

  subcategories <- categories_cn_en %>%
    filter(parent_id %in% input$categories) %>%
    # get parent name (join, remember to rename parent_name field!)
    inner_join(categories_cn_en, by = c("parent_id" = "category_id")) %>%
    # get this name, parent name
    select(Category = category_name_en.x, Parent = category_name_en.y) %>%
    # order by parent then by name
    arrange(Parent, Category)

  subcategories
})
```

### Video Durations

```{r results = "asis"}
renderPlotly({
  plot <- category_data() %>%
    group_by(title, category_name) %>%
    summarise(
      duration = mean(duration, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    ggplot(aes(group=category_name, color=category_name, fill=category_name, x=duration)) +
    geom_density(aes(alpha=ALPHA)) +
    labs(y="density", x="duration (seconds)") +
    scale_color_manual(values = category_colors(), guide="none") +
    scale_fill_manual(values = category_colors(), guide="none") +
    theme_category()
  PLOTLY(plot)
})
```

### Video Exposed Time

```{r results = "asis"}
renderPlotly({
  cd <- category_data()
  plot <- cd %>% ggplot(aes(x=p_hour))
  if (nrow(cd) > 0) {
    plot <- plot +
      # histogram normalized by category_name
      stat_count(aes(y = ..prop.., group = category_name, fill = category_name), position = "dodge") +
      facet_grid(cols = vars(category_name), scales = "fixed") +
      labs(y="factor", x="play hour") +
      scale_color_manual(values = category_colors(), guide="none") +
      scale_fill_manual(values = category_colors(), guide="none") +
      theme_category()
  }
  PLOTLY(plot)
})
```

Engagement
==========
```{css, echo=FALSE}
/* Custom styling for slider to remove blue fill */
.irs-bar,
.irs-bar-edge,
.irs-single, .irs-to, .irs-from {
  background: #ccc !important;
  border-color: #ccc !important;
}
.irs-slider {
  background: #222 !important;
}

/* Additional styling for cleaner appearance */
.chart-title {
  font-size: 16px;
  font-weight: bold;
  text-align: center;
}

/* Reduce padding in controls column */
.section.sidebar {
  padding: 10px;
}
```

```{r setup, include=FALSE}
library(flexdashboard)
library(shiny)
library(ggplot2)
library(dplyr)
library(viridis)
library(readr)

# Load the cleaned data
df <- read_csv("cleaned_engagement_data.csv")

# Ensure proper factor ordering
df <- df %>%
  mutate(plot_tier = factor(plot_tier, 
                           levels = c("first-tier city", "new first-tier city", 
                                    "second-tier city", "third-tier city", "other")))

# Function to get all hours data
get_all_hours_data <- function() {
  df %>%
    select(p_hour, plot_tier, total_engagement) %>%
    arrange(p_hour, plot_tier)
}

# Function to get data for specific hour
get_hour_data <- function(hour) {
  df %>%
    filter(p_hour == hour) %>%
    select(plot_tier, total_engagement) %>%
    arrange(plot_tier)
}

# Function to get summary statistics
get_summary_stats <- function() {
  df %>%
    summarise(
      peak_hour = p_hour[which.max(total_engagement)],
      peak_engagement = max(total_engagement),
      avg_engagement = round(mean(total_engagement)),
      total_records = n(),
      city_count = n_distinct(plot_tier)
    )
}
```

Column {data-width=200}
-----------------------------------------------------------------------

### Controls

```{r}
inputPanel(
  sliderInput(
    "time_hour",
    label = tags$span(style="font-size:13px; font-weight:bold;", "Pick Hour (0-23):"),
    min = 0,
    max = 23,
    value = 0,
    step = 1, 
    round = TRUE,
    sep = "",
    width = "150px",
    animate = animationOptions(interval = 1200, loop = TRUE)
  ),
  
  # Additional control for visualization options
  checkboxInput("show_points", "Show data points", value = TRUE),
  
  # Color scheme selector
  selectInput("color_scheme", "Color Scheme:",
              choices = list("Viridis" = "viridis",
                           "Plasma" = "plasma", 
                           "Inferno" = "inferno",
                           "Magma" = "magma"),
              selected = "viridis"),
  
  # Smoothing option
  checkboxInput("apply_smooth", "Apply smoothing", value = FALSE)
)
```

Column {data-width=600}
-----------------------------------------------------------------------

### Engagement Throughout the Day

```{r}
renderPlot({
  all_data <- get_all_hours_data()
  
  # Base plot
  p <- ggplot(all_data, aes(x = p_hour, y = total_engagement, 
                           color = plot_tier, group = plot_tier)) +
    geom_line(size = 1.2, alpha = 0.8) +
    labs(
      title = "Real User Engagement Throughout the Day by City Tier",
      x = "Hour of Day (0-23)",
      y = "Total Engagement Score",
      color = "City Tier"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 14),
      axis.title = element_text(face = "bold"),
      legend.title = element_text(face = "bold"),
      legend.position = "bottom",
      panel.grid.minor = element_blank()
    ) +
    scale_x_continuous(breaks = seq(0, 23, 4)) +
    geom_vline(xintercept = input$time_hour, color = "red", 
               linetype = "dashed", alpha = 0.7, size = 1)
  
  # Add points if checkbox is selected
  if(input$show_points) {
    p <- p + geom_point(size = 2, alpha = 0.7)
  }
  
  # Add smoothing if selected
  if(input$apply_smooth) {
    p <- p + geom_smooth(method = "loess", se = FALSE, alpha = 0.3)
  }
  
  # Apply color scheme
  if(input$color_scheme == "viridis") {
    p <- p + scale_color_viridis_d()
  } else if(input$color_scheme == "plasma") {
    p <- p + scale_color_viridis_d(option = "plasma")
  } else if(input$color_scheme == "inferno") {
    p <- p + scale_color_viridis_d(option = "inferno")
  } else if(input$color_scheme == "magma") {
    p <- p + scale_color_viridis_d(option = "magma")
  }
  
  p
})
```

### Engagement Heatmap by Hour and City Level

```{r}
renderPlot({
  all_data <- get_all_hours_data()
  
  # Create properly formatted heatmap data
  heatmap_data <- all_data %>%
    mutate(
      hour_formatted = sprintf("%02d", p_hour),
      plot_tier = factor(plot_tier, levels = rev(c("first-tier city", "new first-tier city", 
                                                  "second-tier city", "third-tier city", "other")))
    )
  
  ggplot(heatmap_data, aes(x = hour_formatted, y = plot_tier, fill = total_engagement)) +
    geom_tile(color = "white", size = 0.5) +
    scale_fill_gradient2(low = "lightblue", mid = "orange", high = "red", 
                        midpoint = median(heatmap_data$total_engagement),
                        name = "Total\nEngagement") +
    labs(
      x = "Hour of Day",
      y = "City Level",
      title = "Real Data: Engagement Heatmap by Hour and City Level"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      axis.title.x = element_text(face = "bold"), 
      axis.title.y = element_text(face = "bold"),
      plot.title = element_text(face = "bold", hjust = 0.5, size = 14),
      legend.title = element_text(face = "bold"),
      axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
      axis.text.y = element_text(size = 10),
      panel.grid = element_blank()
    ) +
    # Add vertical line for selected hour
    geom_vline(xintercept = sprintf("%02d", input$time_hour), 
               color = "black", linetype = "dashed", alpha = 0.8, size = 1)
})
```

Column {data-width=400}
-----------------------------------------------------------------------

### Current Hour Analysis

```{r}
renderPlot({
  hour_data <- get_hour_data(input$time_hour)
  
  ggplot(hour_data, aes(x = plot_tier, y = total_engagement, fill = plot_tier)) +
    geom_col(alpha = 0.8, color = "white") +
    labs(
      title = paste("Engagement at Hour", input$time_hour),
      x = "City Tier",
      y = "Total Engagement Score"
    ) +
    theme_minimal(base_size = 12) +
    theme(
      plot.title = element_text(face = "bold", hjust = 0.5, size = 14),
      axis.title = element_text(face = "bold"),
      axis.text.x = element_text(angle = 45, hjust = 1),
      legend.position = "none",
      panel.grid.minor = element_blank()
    ) +
    scale_fill_viridis_d(alpha = 0.8) +
    # Add value labels on bars
    geom_text(aes(label = total_engagement), 
              vjust = -0.5, size = 3, fontface = "bold")
})
```

### Summary Statistics

```{r}
renderTable({
  all_data <- get_all_hours_data()
  hour_data <- get_hour_data(input$time_hour)
  summary_stats <- get_summary_stats()
  
  summary_df <- data.frame(
    Metric = c("Peak Hour", "Peak Engagement", "Current Hour Total", 
               "Average Engagement", "Total Data Points", "City Tiers"),
    Value = c(
      summary_stats$peak_hour,
      format(summary_stats$peak_engagement, big.mark = ","),
      format(sum(hour_data$total_engagement), big.mark = ","),
      format(summary_stats$avg_engagement, big.mark = ","),
      format(summary_stats$total_records, big.mark = ","),
      summary_stats$city_count
    )
  )
  
  summary_df
}, striped = TRUE, hover = TRUE, bordered = TRUE)
```

Chinese Map
===========

```{r}
library(leaflet)

fluidPage(
  tags$head(
    tags$style(HTML("
      .content-wrapper, .right-panel {
        padding-left: 0px !important;
      }
      .map-container {
        position: relative;
        height: 100vh;
        width: 100%;
      }
    "))
  ),
  
  div(class = "map-container",
    # Controls positioned absolutely in top-left
    div(class = "map-controls",
      h4("Map Controls", style = "margin-top: 0;"),
      
      checkboxGroupInput(
        "selected_tiers",
        "Select City Tiers:",
        choices = list(
          "First-tier cities" = "first-tier city",
          "New first-tier cities" = "new first-tier city", 
          "Second-tier cities" = "second-tier city",
          "Third-tier cities" = "third-tier city",
          "Other cities" = "other"
        ),
        selected = c("first-tier city", "new first-tier city", "second-tier city")
      ),
      
      selectInput(
        "map_metric",
        "Display Metric:",
        choices = list(
          "Total Engagement" = "total_engagement",
          "Average Engagement" = "avg_engagement", 
          "User Count" = "user_count",
          "Avg Watch Time" = "avg_watch_time"
        ),
        selected = "total_engagement"
      ),
      
      sliderInput(
        "circle_size_multiplier",
        "Circle Size:",
        min = 0.1,
        max = 3,
        value = 1,
        step = 0.1
      ),
      
      hr(),
      
      # Summary info
      div(id = "map_summary",
        h5("Summary:"),
        textOutput("total_cities"),
        textOutput("total_engagement_map")
      )
    ),
    
    leafletOutput("china_map", width = "100%", height = "100vh")
  )
)
```

```{r}
# Map rendering logic
output$china_map <- renderLeaflet({
  map_data <- get_map_data(input$selected_tiers, input$map_metric)
  
  if (nrow(map_data) == 0) {
    # Empty map if no data
    leaflet() %>%
      addTiles() %>%
      setView(lng = 104.1954, lat = 35.8617, zoom = 4)
  } else {
    # Create color palette based on selected metric
    metric_values <- map_data[[input$map_metric]]
    color_pal <- colorNumeric(
      palette = c("#FFF5F0", "#FB6A4A", "#CB181D", "#67000D"),
      domain = metric_values
    )
    
    # Create size scaling
    size_range <- range(metric_values, na.rm = TRUE)
    size_scale <- function(x) {
      scaled <- (x - size_range[1]) / (size_range[2] - size_range[1])
      (scaled * 20 + 5) * input$circle_size_multiplier
    }
    
    tier_colors <- c(
      "first-tier city" = "#E31A1C",
      "new first-tier city" = "#FF7F00", 
      "second-tier city" = "#1F78B4",
      "third-tier city" = "#33A02C",
      "other" = "#6A3D9A"
    )
    
    leaflet(map_data) %>%
      addTiles() %>%
      setView(lng = 104.1954, lat = 35.8617, zoom = 4) %>%
      addCircleMarkers(
        lng = ~lng,
        lat = ~lat,
        radius = ~size_scale(get(input$map_metric)),
        color = ~tier_colors[plot_tier],
        fillColor = ~color_pal(get(input$map_metric)),
        fillOpacity = 0.7,
        stroke = TRUE,
        weight = 2,
        popup = ~paste0(
          "<b>", fre_city, "</b><br/>",
          "Tier: ", plot_tier, "<br/>",
          "Total Engagement: ", format(total_engagement, big.mark = ","), "<br/>",
          "Avg Engagement: ", round(avg_engagement, 2), "<br/>",
          "Users: ", format(user_count, big.mark = ","), "<br/>",
          "Avg Watch Time: ", round(avg_watch_time, 1), "s"
        ),
        label = ~paste(fre_city, "-", plot_tier)
      ) %>%
      addLegend(
        "bottomright",
        pal = color_pal,
        values = ~get(input$map_metric),
        title = case_when(
          input$map_metric == "total_engagement" ~ "Total Engagement",
          input$map_metric == "avg_engagement" ~ "Avg Engagement",
          input$map_metric == "user_count" ~ "User Count", 
          input$map_metric == "avg_watch_time" ~ "Avg Watch Time (s)",
          TRUE ~ "Value"
        ),
        opacity = 1
      )
  }
})

output$total_cities <- renderText({
  map_data <- get_map_data(input$selected_tiers, input$map_metric)
  paste("Cities displayed:", nrow(map_data))
})

output$total_engagement_map <- renderText({
  map_data <- get_map_data(input$selected_tiers, input$map_metric)
  if (nrow(map_data) > 0) {
    total_eng <- sum(map_data$total_engagement, na.rm = TRUE)
    paste("Total engagement:", format(total_eng, big.mark = ","))
  } else {
    "No data to display"
  }
})
```
