---
title: "WTO Report"
author: "Maciej Zygmanowski (160324), Sofiyan Mohammed (161933)"
date: "2025-05-10"
output:
  html_document:
    theme: cosmo
    css: mystyle.css
    includes:
      before_body: header.html
    toc: true
    toc_float: true
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, results = FALSE, message = FALSE, warning = FALSE)
```

```{r}
library(ggplot2)
library(arrow)
library(tidyr)
library(dplyr)
library(stringr)
library(plotly)
library(readr)
library(fmsb)
library(scales)

GG_THEME = theme_bw() + theme(
  text = element_text(family = "Source Sans Pro")
)
```

```{r csv_to_parquet, eval=FALSE}
# The parquet files are committed anyways so you shouldn't need to run this.
library(parquetize)

csv_to_parquet(
  path_to_file="wto_services_values/services_annual_dataset.csv",
  path_to_parquet="wto_services_values/services_annual_dataset.parquet"
)

csv_to_parquet(
  path_to_file="wto_merchandise_values/merchandise_values_annual_dataset.csv",
  path_to_parquet="wto_merchandise_values/merchandise_values_annual_dataset.parquet"
)
```

```{r load_data, message=TRUE}
services_annual_dataset = arrow::read_parquet("wto_services_values/services_annual_dataset.parquet")
#dplyr::glimpse(services_annual_dataset)
merchandise_values_annual_dataset = arrow::read_parquet("wto_merchandise_values/merchandise_values_annual_dataset.parquet")
#dplyr::glimpse(merchandise_values_annual_dataset)

# concat these tables
dataset = bind_rows(services_annual_dataset, merchandise_values_annual_dataset)

# drop useless columns (having only a single value)
dataset %>% select(-one_of(c("Period", "PeriodCode", "FrequencyCode", "Frequency", "Unit", "UnitCode"))) -> dataset
```

```{r constants}
RP_CODE_WORLD = "000"
CONTINENT_CODES = c("970", "950", "983", "905", "931", "932")

P_CODE_SERVICES_TOTAL = "S"
P_CODE_MERCHANDISE_TOTAL = "TO"

CODE_SERVICES_EXPORT = "ITS_CS_AX5"
CODE_SERVICES_IMPORT = "ITS_CS_AM5"
CODE_SERVICES_EXPORT_FULL = "ITS_CS_AX6"
CODE_SERVICES_IMPORT_FULL = "ITS_CS_AM6"
CODE_MERCHANDISE_EXPORT = "ITS_MTV_AX"
CODE_MERCHANDISE_IMPORT = "ITS_MTV_AM"

CODES_SERVICES = c(CODE_SERVICES_EXPORT_FULL, CODE_SERVICES_IMPORT_FULL)
CODES_MERCHANDISE = c(CODE_MERCHANDISE_EXPORT, CODE_MERCHANDISE_IMPORT)


SHORT_PRODUCT_NAMES = read.csv("./short_product_names.csv")
SHORT_INDICATOR_NAMES = read.csv("./short_indicator_names.csv")

DOLLAR_LABELS=scales::label_dollar(scale_cut = scales::cut_short_scale())
```

```{r preprocess}
# drop non-country Reporters/Partners:
# - 900 are various aggregates like "Europe" and we can filter them out
# - 000 is "world" and we can drop it
# - letter codes are to drop too.
# - The rest are countries and we keep them.
is_country = function(col) {
  is_000 = col == RP_CODE_WORLD
  is_9xx = as.numeric(col) >= 900
  is_9xx[is.na(is_9xx)] = TRUE
  is_9xx
  return(!(is_000 | is_9xx))
}

dataset[is_country(dataset$ReporterCode) & is_country(dataset$PartnerCode),] -> dataset_only_countries

# List of product codes
product_codes = dataset %>% select(ProductCode,Product) %>% unique
#print(product_codes, n=1000)

# List of top-level product codes
product_codes %>% filter((nchar(ProductCode) == 2) & (ProductCode != "TO")) -> top_level_product_codes

max_year = max(dataset$Year)
```

<!----- HERE ACTUAL START OF THE REPORT ----->

# Summary

This report shows exploratory data analysis of a 2023 WTO report of import and export of countries/groups of countries in various sectors of their economy. Some interesting findings (concerning, for example, relation of USA with Ireland) are presented.

# Analysis

## Top 5 countries by services/merchandise import/export

On this plot, we see top 5 countries (for year `r max_year`) for every indicator (import/export
of services/merchandise).

We can see that:

- USA is both top importer and exporter of services, and also top importer
  of merchandise
- China is top exporter of merchandise

Recessions are marked <span style="background-color:pink">pink</span>. The 2008-2009 one is quite visible.
Surprisingly, the 1980 recession has a bump.

```{r}
show_over_time = function() {
  # dataset where one side is World, and the other is country
  INDICATORS = c(CODE_SERVICES_EXPORT, CODE_SERVICES_IMPORT, CODE_MERCHANDISE_EXPORT, CODE_MERCHANDISE_IMPORT)
  should_keep_code = function(df) {
    indicator = df$IndicatorCode %in% INDICATORS
    reporter = df$ReporterCode
    partner = df$PartnerCode
    reporter_is_country = is_country(reporter)
    partner_is_country = is_country(partner)
    return(indicator & ((reporter == RP_CODE_WORLD & partner_is_country) | (partner == RP_CODE_WORLD & reporter_is_country)))
  }
  
  dataset_world_to_country <- dataset[should_keep_code(dataset), ]
  
  top_reporters_total = dataset_world_to_country %>%
    group_by(ReporterCode,Reporter,IndicatorCode,Indicator) %>%
    summarise(sum_value = sum(Value)) %>%
    group_by(IndicatorCode,Indicator) %>%
    slice_max(order_by = sum_value, n = 5)
  
  reporters_by_year = dataset_world_to_country %>%
    group_by(ReporterCode,Reporter,IndicatorCode,Year) %>%
    summarise(sum_value = sum(Value)) %>%
    inner_join(top_reporters_total, c("ReporterCode", "Reporter", "IndicatorCode")) %>%
    inner_join(SHORT_INDICATOR_NAMES, by="IndicatorCode")
  
  # thx https://stackoverflow.com/a/41637932
  Reces_table <-read.csv("./recessions.csv")
  Reces_table$Start <- parse_date(Reces_table$Start, format="%Y-%m-%d")
  Reces_table$End <- parse_date(Reces_table$End, format="%Y-%m-%d")
  
  grid_vector = c()
  indicator_codes = 
    for (i in 1:length(INDICATORS)) {
      indicator <- INDICATORS[i]
      reporters_by_year_for_indicator <- reporters_by_year %>% filter(IndicatorCode == indicator)
      reporters_by_year_for_indicator$Year = parse_date(as.character(reporters_by_year_for_indicator$Year), format="%Y") 
      print(reporters_by_year_for_indicator)
      # Create the bar plot

      print(min(reporters_by_year_for_indicator$Year))
      print(max(reporters_by_year_for_indicator$Year))
      print(Reces_table)
      
      p <- ggplot(reporters_by_year_for_indicator) + GG_THEME +
              geom_line(aes(color=Reporter, x=Year, y=sum_value.x*1e6)) +
              scale_y_log10(labels=DOLLAR_LABELS) +
              geom_rect(data = Reces_table, inherit.aes = FALSE,
                        aes(xmin=Start, xmax=End, ymin=0, ymax=+Inf), 
                        fill='pink', alpha=0.5) +
              labs(title=reporters_by_year_for_indicator$IndicatorShortName, y="Value (mln $)")
              
      # append plot
      print(p)
      #grid_vector <- c(grid_vector, list(p))
    }

  # Create a grid layout
  #do.call(gridExtra::grid.arrange, c(grid_vector, ncol = 2))
}
show_over_time()
```

<!--
## Import/Export of selected countries by product

```{r}
countries = c("156", "840", "616") # China, USA, Poland

dataset %>%
  filter((ReporterCode %in% countries) &
           (ProductCode %in% top_level_product_codes$ProductCode) &
           (Year == max_year) & (PartnerCode == RP_CODE_WORLD)) ->
  export_by_sector

export_by_sector

EXPORT=c(CODE_MERCHANDISE_EXPORT,CODE_SERVICES_EXPORT_FULL)
IMPORT=c(CODE_MERCHANDISE_IMPORT,CODE_SERVICES_IMPORT_FULL)
export_by_sector <- export_by_sector %>%
  mutate(ImportExport=ifelse(IndicatorCode %in% EXPORT, "Export",
                      ifelse(IndicatorCode %in% IMPORT, "Import", NA)))

ggplot(export_by_sector) + GG_THEME +
  geom_bar(stat="identity", aes(x=str_c(ProductCode," ",Product), y=Value)) +
  facet_grid(ImportExport~Reporter) +
  coord_flip()
```
-->

## Top 5 services sectors per continent

```{r}
plot_top_5_sectors_per_continent = function(IE) {
  dataset_continents <- dataset %>% filter(
    (ReporterCode %in% CONTINENT_CODES) & (PartnerCode == RP_CODE_WORLD) &
      (Year == max_year) & (IndicatorCode %in% IE) &
      (ProductCode %in% top_level_product_codes$ProductCode)
  ) %>% group_by(ReporterCode) %>% slice_max(order_by=Value, n = 4) %>% ungroup() %>%
    mutate(Reporter = recode(Reporter, !!!c("Australia and New Zealand"="Australia and N.Zeal.", "South and Central America and the Caribbean"="South America")))

  grid_vector = c()
  for (i in 1:length(CONTINENT_CODES)) {
    continent_code <- CONTINENT_CODES[i]
    continent_data <- dataset_continents %>% filter((ReporterCode == continent_code)) %>%
      inner_join(SHORT_PRODUCT_NAMES, by="ProductCode")
    continent_name = continent_data$Reporter
    
    # Create the bar plot
    p <- ggplot(continent_data, aes(x = ProductShortName, y = Value*1e6)) + GG_THEME +
      geom_bar(stat = "identity") +
      labs(title = continent_name,
           y = "Value (mln $)",x = "") +
      scale_y_continuous(labels=DOLLAR_LABELS) +
      theme(axis.text.x = element_text(angle = 30, hjust=1), plot.title = element_text(size=10))
      
    
    # append plot
    grid_vector <- c(grid_vector, list(p))
  }
  
  # Create a grid layout
  do.call(gridExtra::grid.arrange, c(grid_vector, ncol = 3))
}
```

### Export

```{r}
plot_top_5_sectors_per_continent(EXPORT)
```

Most continents export travel, it seems. Especially Australia.

### Import

```{r}
plot_top_5_sectors_per_continent(IMPORT)
```

### Interesting findings

"IP Use Charges"??? let's dive in what specifically they are doing...

```{r, results='markup'}
IP_USE_CHARGES = "SH"
dataset %>% filter(
  str_detect(ProductCode, IP_USE_CHARGES) &
    is_country(dataset$ReporterCode) &
    is_country(dataset$PartnerCode) &
    (nchar(ProductCode) == 3) &
    (Year == max_year) &
    (IndicatorCode %in% EXPORT)
) %>%
  arrange(-Value) %>%
  select(Reporter,Partner,Product,Value) %>%
  lapply(., iconv, to = "UTF-8") %>% tibble::as_tibble() %>%
  head(20) %>%
  knitr::kable(format="html")
```

---

Export from US to Ireland is interesting here... let's see when that bump
happened. On this plot, Ireland is compared to another EU country with similar
import/export values (Germany):

```{r}
dataset %>% filter(
  (ProductCode == "S") &
  (dataset$Reporter %in% c("Ireland", "Germany")) &
  (dataset$PartnerCode == "840") &
  (IndicatorCode %in% c(CODE_SERVICES_EXPORT_FULL, CODE_SERVICES_IMPORT_FULL))
) -> double_irish

double_irish <- double_irish %>%
  mutate(ImportExport=ifelse(IndicatorCode %in% EXPORT, "Export",
                             ifelse(IndicatorCode %in% IMPORT, "Import", NA)))

double_irish %>% select(ProductCode)

double_irish %>% ggplot(aes(x=Year, y=Value, color=Reporter, linetype=ImportExport)) + GG_THEME +
  geom_line() +
  geom_point() +
  labs(title="Ireland Import/Export of Services from/to USA over time", linetype="Import/Export", y="Value (mln $)")
```

This is how it looks when split by sector:

```{r results='markup'}
dataset %>% filter(
  (ProductCode %in% top_level_product_codes$ProductCode) &
    (dataset$Reporter == "Ireland") &
    (dataset$PartnerCode == "840") &
    (IndicatorCode %in% c(CODE_SERVICES_IMPORT_FULL))
) %>% inner_join(SHORT_PRODUCT_NAMES, by="ProductCode") -> double_irish

double_irish <- double_irish %>%
  mutate(ImportExport=ifelse(IndicatorCode %in% EXPORT, "Export",
                             ifelse(IndicatorCode %in% IMPORT, "Import", NA)))

double_irish <- double_irish %>%
  mutate(ProductShortName=ifelse(ProductCode == "SH", ProductShortName, "Other")) %>%
  group_by(Year,ProductShortName) %>%
  summarize(Value=sum(Value))

double_irish %>% ggplot(aes(x=Year, y=Value, fill=ProductShortName)) + GG_THEME +
           geom_area(position="stack") +
           labs(title="Ireland Import of Services from/to USA over time, by sector",
                y="Value (mln $)", fill="Product")
```

We can clearly see that "IP Use Charges" rises sharply somewhere between 2016
and 2020 (there is no data points there, so we can't be sure when exactly).

## Compare how import correlates with export in various economies

```{r results='markup'}
# Import/export per country in continent
# X-Services Import
# Y-Services Export
# Time-Year
# Color-Continent

country_continents = read.csv("country_continents.csv")
country_continents$CountryCode = as.character(country_continents$CountryCode)

ie_per_country_base <- dataset %>%
  filter(PartnerCode == RP_CODE_WORLD & ((ProductCode == P_CODE_SERVICES_TOTAL) | (ProductCode == P_CODE_MERCHANDISE_TOTAL))) %>%
  mutate(ImportExport=ifelse(IndicatorCode %in% EXPORT, "Export", ifelse(IndicatorCode %in% IMPORT, "Import", NA)))

ie_per_country <- ie_per_country_base %>%
  pivot_wider(id_cols=c(ReporterCode,Reporter,Year,ProductCode), names_from=ImportExport, values_from=Value) %>%
  inner_join(country_continents, by=c("ReporterCode"="CountryCode"), relationship="many-to-many") %>%
  filter(Year==max_year) %>%
  mutate(ProductCode = recode(ProductCode, !!!c("S" = "Services", "TO" = "Merchandise")), Import=Import*1e6, Export=Export*1e6)

options(scipen = 999)

breaks <- 10^(5:13)

plot <- ggplot(ie_per_country, aes(x=Import,y=Export,color=Continent,text=CountryName)) +
  geom_abline(slope=1, intercept=0) +
  geom_point() +
  geom_smooth(inherit.aes=F, aes(x=Import,y=Export)) +
  labs(y="Export", x="Import") +
  facet_grid(~ProductCode) +
  scale_x_log10(breaks=breaks, labels=DOLLAR_LABELS) +
  scale_y_log10(breaks=breaks, labels=DOLLAR_LABELS)

ggplotly(plot, tooltip=c("text"))
```

Import and export compared. **Black** line is the ideal, expected situation
where import equals export. The <span style="color:blue">**blue**</span> smooth
line shows how the correlation shows up in reality - smaller economies
are skewed towards importing, especially for merchandise.

We can also see that Asia & Europe generally import & export more than other
continents, due to being more developed. North America is an interesting special
case, with most countries being on the low end except some outliers (e.g USA).
Africa and South America are in the middle. On the opposite side, we see
Oceania, where most countries are small islands with not much opportunity to
develop, and, because of being isolated, they have to import products
from continental countries.

<!--
## Advertising exports

(only USA, only to some countries)

(just for fun)

```{r}
PRODUCT_ADVERTISING = "SJ221"

dataset_only_countries %>%
  filter((ProductCode == PRODUCT_ADVERTISING) & (Year == max_year)) %>%
  # THIS lapply FIXES AN ERROR:
  #Error in grid.Call(C_textBounds, as.graphicsAnnot(x$label), x$x, x$y,  : 
  #                     invalid string in PangoCairo_Text
  lapply(., iconv, to = "UTF-8") %>% tibble::as_tibble() %>%
  mutate(Value=as.numeric(Value)) %>%
  ggplot() + GG_THEME +
  geom_bar(stat="identity", aes(x=str_c(PartnerCode, " ", Partner),y=Value)) +
  coord_flip()
```
-->

## Merchandise profile for top countries

On both plots, value is % of the largest value.

### Export

```{r}
# Short name mapping for Product
short_names <- c(
  "Agricultural products" = "Agric.",
  "Automotive products" = "Auto",
  "Transport equipment" = "Transport",
  "Total merchandise" = "Total",
  "Chemicals" = "Chemicals",
  "Clothing" = "Clothing",
  "Electronic data processing and office equipment" = "ElecData",
  "Food" = "Food",
  "Fuels" = "Fuels",
  "Fuels and mining products" = "FuelsMin",
  "Integrated circuits and electronic components" = "IC/Elec",
  "Iron and steel" = "IronSteel",
  "Machinery and transport equipment" = "Machinery",
  "Manufactures" = "Manuf.",
  "Office and telecom equipment" = "OfficeTel",
  "Pharmaceuticals" = "Pharma",
  "Telecommunications equipment" = "Telecom",
  "Textiles" = "Textiles"
)

# --- Top countries
exports <- merchandise_values_annual_dataset %>%
  filter(
    IndicatorCode == "ITS_MTV_AX",
    Partner == "World",
    ProductCode != "TO"                    # <-- Added filter here
  )

country_coverage <- exports %>%
  group_by(Reporter) %>%
  summarize(n_rows = n())
top_countries <- country_coverage %>% arrange(desc(n_rows)) %>% head(5)
selected_countries <- c("Australia", "Canada", "Japan")  # Use last plot's countries

# --- Radar data with short names
radar_df <- exports %>%
  filter(Reporter %in% selected_countries, Year == max_year, ProductCode != "TO") %>%    # <-- Added filter here
  group_by(Reporter, Product) %>%
  summarize(Value = sum(Value, na.rm=TRUE)) %>%
  group_by(Reporter) %>%
  mutate(Percent = Value / sum(Value) * 100) %>%
  ungroup() %>%
  mutate(Product_short = recode(Product, !!!short_names)) %>%
  select(Reporter, Product_short, Percent)

# Reshape for radar
radar_wide <- radar_df %>%
  tidyr::pivot_wider(names_from = Product_short, values_from = Percent, values_fill = 0)

max_row <- radar_wide %>% select(-Reporter) %>% summarise_all(~max(.))
min_row <- radar_wide %>% select(-Reporter) %>% summarise_all(~min(.))
radar_chart_data <- as.data.frame(rbind(max_row, min_row, radar_wide %>% select(-Reporter)))
rownames(radar_chart_data) <- c("Max", "Min", radar_wide$Reporter)

# --- Make less-intrusive axis color
percent_col <- "#2e4053"  # Changed to requested color

# Plot
country_colors <- c("blue", "orange", "forestgreen")[1:length(selected_countries)]

radarchart(
  radar_chart_data,
  axistype=1,
  pcol=country_colors,
  pfcol=adjustcolor(country_colors, alpha.f=0.22),
  plwd=3,
  plty=c(1,2,3),
  title=paste0("Merchandise Exports Profile (", max_year, ")"),
  cglty=1,
  cglcol=percent_col,           # axis grid lines in requested color
  axislabcol=percent_col,       # axis (% lines) labels in requested color
  vlcex=0.95
)
legend("bottomright", legend = selected_countries, col = country_colors, lwd=3, lty=1:3, bg="white")
```

### Import

```{r}
imports <- merchandise_values_annual_dataset %>%
  filter(
    IndicatorCode == "ITS_MTV_AM",    # <--- Imports
    Partner == "World",
    ProductCode != "TO"
  )

country_coverage <- imports %>%
  group_by(Reporter) %>%
  summarize(n_rows = n())
top_countries <- country_coverage %>% arrange(desc(n_rows)) %>% head(5)
selected_countries <- c("Australia", "Canada", "Japan")  # Use last plot's countries

# --- Radar data with short names
radar_df <- imports %>%
  filter(Reporter %in% selected_countries, Year == max_year, ProductCode != "TO") %>%    # <-- Added filter here
  group_by(Reporter, Product) %>%
  summarize(Value = sum(Value, na.rm=TRUE)) %>%
  group_by(Reporter) %>%
  mutate(Percent = Value / sum(Value) * 100) %>%
  ungroup() %>%
  mutate(Product_short = recode(Product, !!!short_names)) %>%
  select(Reporter, Product_short, Percent)

# Reshape for radar
radar_wide <- radar_df %>%
  tidyr::pivot_wider(names_from = Product_short, values_from = Percent, values_fill = 0)

max_row <- radar_wide %>% select(-Reporter) %>% summarise_all(~max(.))
min_row <- radar_wide %>% select(-Reporter) %>% summarise_all(~min(.))
radar_chart_data <- as.data.frame(rbind(max_row, min_row, radar_wide %>% select(-Reporter)))
rownames(radar_chart_data) <- c("Max", "Min", radar_wide$Reporter)

# --- Make less-intrusive axis color
percent_col <- "#2e4053"  # Changed to requested color

# Plot
country_colors <- c("blue", "orange", "forestgreen")[1:length(selected_countries)]

radarchart(
  radar_chart_data,
  axistype=1,
  pcol=country_colors,
  pfcol=adjustcolor(country_colors, alpha.f=0.22),
  plwd=3,
  plty=c(1,2,3),
  title=paste0("Merchandise Imports Profile (", max_year, ")"),
  cglty=1,
  cglcol=percent_col,           # axis grid lines in requested color
  axislabcol=percent_col,       # axis (% lines) labels in requested color
  vlcex=0.95
)
legend("bottomright", legend = selected_countries, col = country_colors, lwd=3, lty=1:3, bg="white")
```
